### 2019-10-13 MSB1015_Assignment2_SuzannetenHage.Rmd
Improved documentation on section 2
Some description of which descriptors were used was missing, so this has been added.

---
title: "MSB1015_Assignment 2"
author: "Suzanne ten Hage"
date: "12 october 2019"
output: html_notebook
---

**Project Description**
Chemical properties, such as the boiling point, can be derived from the structure of a chemical compound. In 1947, Harry Wiener already made a correlation model to link structural features to boiling points (ref1). The idea to use mathematical models to predict chemical properties from compound structures has been expaned since then.

In this project I use a SPARQL query to obtain the smiles and boiling points of simple alkanes from WikiData (ref2). I use the smiles to get descriptors from the chemical development kit (CDK) database (ref3-6). These descriptors contain information on the structural properties of the alkanes (see section 2 for more details). Finally, I train a Partial Least Squares (PLS) model to predict these properties from the chemical properties of the compounds and plot the results.

**0. Installation**
The project requires several packages. The piece of code in this section checks automatically for missing packages and installs them.

The rJava package requires Java to be installed. The code has been developed using Java version 1.8.0_191. A tutorial on how to install this Java Version can be found here: https://downlinko.com/download-install-jdk-8-windows.html

```{r}
#Required packages for this code. 
packages <- c("WikidataQueryServiceR", 
              "rJava", 
              "rcdk", 
              "caTools",
              "pls", 
              "Metrics"
              )

#Installation of missing packages 
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
    install.packages(setdiff(packages, rownames(installed.packages()))) 
    }

#Load packages
library("WikidataQueryServiceR")
library('rJava')
library('rcdk')
library('caTools')
library('pls')
library('Metrics')

```

**1. Get the data from WikiData.** 
In this section I use the WikidataQueryServiceR package to run a SPARQL query that obtains all simple alkines with known boiling points and smiles from WikiData.

```{r}

#The function query_wikidata runs the SPARQL query to obtain data from the Wikidata database. This returns a dataframe containing the alkane names, boiling points, units and smiles. It also contains the wikidata-links for compounds and units. There are 134 alkanes at the time of writing.

data_wikidata = query_wikidata('SELECT DISTINCT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
    ?comp wdt:P31/wdt:P279* wd:Q41581 ;
    p:P2102 [ps:P2102 ?bp ;           
             psv:P2102/wikibase:quantityUnit  
             ?bpUnit         
            ] ;
    wdt:P233 ?smiles.
    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } 
    } '
)

```
Description of WikiData numbers:
* P31 = instance of
* P279 = subclass of
* Q41581 = alkane
* P2102 = boiling point
* P233 = canonical smiles

The SPARQL query can easily be changed to obtain different data. For example, one could get the smiles of alkynes instead of alkanes by changing Q41581 into Q159226. Look at https://www.wikidata.org/wiki/Wikidata:Main_Page to find the right numbers.

**2. Get the descriptors of the alkane structures**
In this section, I use four descriptors for all alkanes from CDK. I can obtain the descriptors based on the alkane smiles. These descriptors will be used later as predictors in the PLS model.

*APolDescriptor:*
This describes the sum of the atomic polarizabilities.

*WienerNumbersDescriptor:*
This gives the Wiener numbers (see ref1). It returns the Wiener Path Number and the Wiener Polarity Number.

*MDEDescriptor:*
Returns the Molecular Distance Edge.

*FragmentComplexityDescriptor:*
Returns the complexity of a system.

```{r}

#Reformatting of smiles into correct format for CDK. Kekulise checks for electrons, and prevents parsing of incorrect smiles. 
parsed_smiles <- parse.smiles(data_wikidata$smiles, kekulise=TRUE)

#Determine which descriptors are of interest. To add/remove descriptors, change this vector to include the desired descriptors. 
descriptor_names <- c(
'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.MDEDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.FragmentComplexityDescriptor'
)

#Get descriptor values form the CDK database. 
data_descriptors <- eval.desc(parsed_smiles, descriptor_names)

#At the time of writing, this returns a dataframe of 23 variables describing 134 alkanes. 
```

**3. Data pre-processing**
3a) Not all boiling points in the database are in kelvin, which is the SI unit for temperature. Before I continue, I make sure all temperatures are in kelvin.

```{r}

##CELCIUS -> KELVIN
#Find all rownumbers of the values in Celcius. 
celcius <-  which(grepl('degree Celsius', data_wikidata$bpUnitLabel))

for (i in 1:length(celcius)) {
  #Convert the value to kelvin
  data_wikidata$bp[celcius[i]] <- data_wikidata$bp[celcius[i]] + 273.15
  #Convert the label to kelvin. 
  data_wikidata$bpUnitLabel[celcius[i]] <- 'kelvin'
}

##FAHRENHEIT -> KELVIN
#Find all rownumbers of the values in Fahrenheit. 
fahrenheit <- which(grepl('degree Fahrenheit',data_wikidata$bpUnitLabel))

for (i in 1:length(fahrenheit)) {
  #Convert the value to kelvin
  data_wikidata$bp[fahrenheit[i]] <- (data_wikidata$bp[fahrenheit[i]] + 459.67) * (5/9)
  #Convert the label to kelvin. 
  data_wikidata$bpUnitLabel[fahrenheit[i]] <- 'kelvin'
}

```

3b) Combine the descriptors and boiling points in one dataset, so we can use this dataset for the PLS. I remove the columns with (almost) all zeros, because 
```{r}

#Fuse the boiling points from data.raw and descriptors into one dataset. 
data_combined <- data.frame(data_descriptors, data_wikidata$bp)

#Define which columns to remove. 
remove_columns <- c("MDEC.13", "MDEC.14", "MDEC.22", "MDEC.23", "MDEC.24", "MDEC.33", "MDEC.34", "MDEC.44", "MDEO.11", "MDEO.12", "MDEO.22", "MDEN.11", "MDEN.12", "MDEN.13", "MDEN.22", "MDEN.23", "MDEN.33")

#Remove all empty columns from the dataset.
data_combined <- data_combined[ , -which(names(data_combined) %in% remove_columns)]

```

3c) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data. In the following code you can easily change the size of the test data by adjusting percentage_test.

```{r}

#Percentage of data to go into the test-set. 
percentage_test <- 0.25

#Splitting data into a training and test set. 
set.seed(101) 
sample = sample.split(data, SplitRatio = percentage_test)
train = subset(data, sample == FALSE)
test  = subset(data, sample == TRUE)

```

**4. Null model**
The null model is the most simple model possible: the mean of the data. This model can be used as a reference to see whether the PLS model performs better.

```{r}

#Create null_model
null_model <- mean(train$data_raw.bp)

#Get error of null model
error_null_model <- rmse(test$data_raw.bp, null_model)

```

**5. Partial Least Squares**
Here, I train the PLS model with the training data, and use the model to predict the boiling points of the test data.

```{r}

#Training of the PLS model. 
PLS_model <- plsr(data_raw.bp ~ ., 
                  data = train, 
                  validation = "CV"
                  )

#test.try <- test[,c(1:34)];
#Predict the boiling points of the test set. 
predicted.bp <- predict(PLS_model, test)

#Calculate the error. 
error <- RMSEP(PLS_model, newdata = test)

```

**6. Visualization**

```{r}

#Plot the error vs. the number of components
plot(error)

#Plot the predicted data vs. the original data. 
plot(test$data_raw.bp, predicted.bp[,,3], xlim=c(300,1000), ylim=c(300,1000))

plot(test$data_raw.bp, rep(null_model, length(test$data_raw.bp)))

#Correlation between predicted and original test data. 
corcoef <- cor(predicted.bp[,,3], test$data_raw.bp)

```

**7. References**
ref1: Wiener H. Structural Determination of Paraffin Boiling Points. Journal of the American Chemical Society. 1947 Jan;69(1):17–20.
ref2: https://www.wikidata.org/wiki/Wikidata:Main_Page (12-10-2019)
ref3: Willighagen et al. The Chemistry Development Kit (CDK) v2.0: atom typing, depiction, molecular formulas, and substructure searching. J. Cheminform. 2017; 9(3), doi:10.1186/s13321-017-0220-4
ref4: May and Steinbeck. Efficient ring perception for the Chemistry Development Kit. J. Cheminform. 2014, doi:10.1186/1758-2946-6-3
ref5: Steinbeck et al. Recent Developments of the Chemistry Development Kit (CDK) - An Open-Source Java Library for Chemo- and Bioinformatics. Curr. Pharm. Des. 2006; 12(17):2111-2120, doi:10.2174/138161206777585274
ref6: Steinbeck et al. The Chemistry Development Kit (CDK): An Open-Source Java Library for Chemo- and Bioinformatics. J. Chem. Inf. Comput. Sci. 2003 Mar-Apr; 43(2):493-500, doi:10.1021/ci025584y